---
title: "MA615 Buoy Project"
---
1.Introduction:
The project is to analyze whether there is an evidence of global warming in the data collected by the single weather buoy in the NOAA National Data buoy center. This can be achieved by deriving empirical relationships (regression models) between years and two measurements: air temperature(ATMP) and wave temperature(WTMP).

2.Dataset
2.1 Import the 20-year data from the website.
```{r}
library(tidyverse)
library(stringr)

url_1 <- "http://www.ndbc.noaa.gov/view_text_file.php?filename=mlrf1h"
url_2 <- ".txt.gz&dir=data/historical/stdmet/"
years <- c(1999:2018)
urls <- str_c(url_1, years, url_2, sep = "")
filenames <- str_c("mr", years, sep = "")

# Year 1999 - 2006
for(i in 1:8){
  suppressMessages(
    # Fill any missing values with NA:
    assign(filenames[i], read.table(urls[i], header = TRUE, fill = TRUE))
  )
  
}
# Year 2007 - 2018
for(i in 9:20){
  suppressMessages(
    # Fill any missing values with NA and use the same column names as year 2006
    assign(filenames[i], read.table(urls[i], header = FALSE, 
                                    fill = TRUE, col.names = colnames(mr2006))),
  )
  
}

mr1999$TIDE <- NA
n <- length(urls)
for (i in 1:n){
  file <- get(filenames[i])
  colnames(file)[1] <-"YYYY"
  if(ncol(file) == 18){
    file <- subset(file, select = -mm )
  }
  if(i == 1){
    MR <- file
  }else{
    MR <- rbind.data.frame(MR, file)
  }
  
}
```

2.2 Tidy Data
```{r}
# delete columns in MR
MR_1<-MR[,-c(5:12)]
View(MR_1)
MR_2<-MR_1[,-c(7:9)]
(MR_2)

#replace all 999 values in column ATMP and column WTMP with NA 
MR_2[MR_2==999]<-NA
MR_2
```

```{r}
#transform the date-time data into posix numbers using lubridate.
library(lubridate)
MR_2$date=make_datetime(MR$YYYY,MR$MM,MR$DD,MR$hh)
MR_3<-transform(MR_2,date=MR_2$date)
(MR_3)
```
```{r}
#change the order of columns.
MR_4=select(MR_3,7,everything())
MR_4
```

```{r}
#Pick the 12:00 data on each day.
MR_5=subset(MR_4, grepl("12", MR_4$hh))
MR_5
```

3.Models
This step is to build linear regression models on analyzing relationships between: 1.the average air temperature and years. 2.the average wave temperature and years. 3.the maximum air temperature and years. 4.the maximum wave temperature and years. 5.the minimum air temperature and years. 6.the minimum wave temperature and years.
```{r}
#Calculte the means of ATMP and WTMP.
library(dplyr,warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
MR_mean=MR_5 %>% group_by(YYYY) %>% summarise(mean_ATMP =  mean(ATMP,na.rm=T), mean_WTMP=mean(WTMP,na.rm=T))
MR_mean
```
```{r}
library(ggplot2)
meanplot<-ggplot(data = MR_mean) +geom_point(mapping = aes(x =YYYY, y = mean_ATMP),color="black")+geom_point(mapping = aes(x =YYYY, y = mean_WTMP),color="blue")
meanplot+labs(x="Year",y="Temperature")
```
The 20 blue points represent the average wave temperature for each year from year 1999 to 2018 and all black points represent the average air temperature for each year. It is extremely clear from the plot that there is no overlap between the average air temperature and the average wave temperature in each year. Besides, the plot shows that blue points and black points move in a similar pattern so the average wave temperature and the average air temperature might both are evidences of global warming or not. From 2015 to 2018, the average air temperature and the average wave temperature are generally decreasing and they might be increasing after year 2018 based on the pattern shown in the plot. 

```{r}
library("rstanarm")
M1<-stan_glm(mean_ATMP~YYYY,data=MR_mean,refresh=0)
coef(M1)
```
```{r}
label = paste("y = ", round(coef(M1)[1],4), " + ", round(coef(M1)[2],4), "x", sep="")
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
plot(MR_mean$YYYY,MR_mean$mean_ATMP,pch=20,xlab="Year",ylab="Average Air Temperature")
abline(coef(M1))
text(2015,24.5,label)
```
Model 1: fit a linear regression model on the average air temperature and years. Points are scattered.
```{r}
M2<-stan_glm(mean_WTMP~YYYY,data=MR_mean,refresh=0)
coef(M2)
```
```{r}
label = paste("y = ", round(coef(M2)[1],4), " + ", round(coef(M2)[2],4), "x", sep="")
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
plot(MR_mean$YYYY,MR_mean$mean_WTMP,pch=20,xlab="Year",ylab="Average Wave Temperature")
abline(coef(M2))
text(2015,26.5,label)
```
Model 2: fit a linear regression model on the average wave temperature and years. Points are scattered.
Based on Model 1 and Model 2, the two regression lines have nearly the same coefficients for year with a small difference in intercepts. From two models, as the year increases, the average air temperature and the average wave temperature both increases.

```{r}
#Calulate maximums of ATMP and WTMP.
MR_MAX=MR_5 %>% group_by(YYYY) %>% summarise(MAX_ATMP =  max(ATMP,na.rm=T), MAX_WTMP =  max(WTMP,na.rm=T))
MR_MAX
```
```{r}
maxplot<-ggplot(data = MR_MAX) +geom_point(mapping = aes(x =YYYY, y = MAX_ATMP),color="black")+geom_point(mapping = aes(x =YYYY, y = MAX_WTMP),color="blue")
maxplot+labs(x="Year",y="Temperature")
```
The 20 blue points represent the maximum wave temperature for each year from year 1999 to 2018 and all black points represent the maximum air temperature for each year. The points of maximum air temperatures and the maximum wave temperatures are scattered. From 2017 to 2018, the maximum air temperatures are exactly the same and the maximum wave temperatures are nearly the same. 
```{r}
M3<-stan_glm(MAX_ATMP~YYYY,data=MR_MAX,refresh=0)
coef(M3)
```
```{r}
label = paste("y = ", round(coef(M3)[1],2), " + ", round(coef(M3)[2],2), "x", sep="")
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
plot(MR_MAX$YYYY,MR_MAX$MAX_ATMP,pch=20,xlab="Year",ylab="Maximum Air Temperature")
abline(coef(M3))
text(2015,29.7,label)
```
Model 3: fit a linear regression model on the maximum air temperature and years. Points are scattered.
```{r}
M4<-stan_glm(MAX_WTMP~YYYY,data=MR_MAX,refresh=0)
coef(M4)
```

```{r}
label = paste("y = ", round(coef(M4)[1],4), " + ", round(coef(M4)[2],4), "x", sep="")
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
plot(MR_MAX$YYYY,MR_MAX$MAX_WTMP,pch=20,xlab="Year",ylab="Maximum Wave Temperature")
abline(coef(M4))
text(2015,30.4,label)
```
Model 4: fit a linear regression model on the maximum wave temperature and years. Based on model 3 and model 4, as the year increases one year, the maximum air temperature will change more compared with the maximum wave temperature.
```{r}
#Calculate minimums of ATMP and WTMP.
library(dplyr,warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
MR_min=MR_5 %>% group_by(YYYY) %>% summarise(min_ATMP=min(ATMP,na.rm=T), min_WTMP=min(WTMP,na.rm=T))
MR_min
```
```{r}
minplot<-ggplot(data = MR_min) +geom_point(mapping = aes(x =YYYY, y = min_ATMP),color="black")+geom_point(mapping = aes(x =YYYY, y = min_WTMP),color="blue")
minplot+labs(x="Year",y="Temperature")
```
The 20 blue points represent the minimum wave temperature for each year from year 1999 to 2018 and all black points represent the minimum air temperature for each year. No overlap between these blue points and black points. The minimum air temperatures have bigger differences between years.
```{r}
M5<-stan_glm(min_ATMP~YYYY,data=MR_min,refresh=0)
label = paste("y = ", round(coef(M5)[1],2), " + ", round(coef(M5)[2],2), "x", sep="")
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
plot(MR_min$YYYY,MR_min$min_ATMP,pch=20,xlab="Year",ylab="Minimum Air Temperature")
abline(coef(M5))
text(2015,8,label)
```
Model 5: fit a linear regression model on the minimum air temperature and years. 
```{r}
M6<-stan_glm(min_WTMP~YYYY,data=MR_min,refresh=0)
label = paste("y = ", round(coef(M6)[1],2), " + ", round(coef(M6)[2],2), "x", sep="")
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
plot(MR_min$YYYY,MR_min$min_WTMP,pch=20,xlab="Year",ylab="Minimum Wave Temperature")
abline(coef(M6))
text(2015,21,label)
```
Model 6: fit a linear regression model on the minimum wave temperature and years. Based on model 5 and model 6, as the year increases one year, the minimum air temperature will change more compared with the minimum wave temperature.

4.Result
From linear regression models in the part 3, we can conclude that there is an evidence of global warming in the data. As observed from these models, the coefficients for year are positive. Therefore, as the year increases one year, the means, minimums and maximums of air temperatures and wave temperatures are expected to increase, which shows that the global warming is occuring.

References：
U.S Dept of Commerce (2020) National Oceanic and Atmospheric Administration's National Data Buoy Center Center of Excellence in Marine Technology. Available at: https://www.ndbc.noaa.gov/station_page.php?station=44013 (Accessed: Sep 25 2020).

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686

Hadley Wickham (2019). stringr: Simple, Consistent Wrappers for Common String Operations. R package version 1.4.0.
  https://CRAN.R-project.org/package=stringr
  
Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3), 1-25. URL
  http://www.jstatsoft.org/v40/i03/. 
  
Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr: A Grammar of Data Manipulation. R package version 1.0.2.
  https://CRAN.R-project.org/package=dplyr
  
H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
 
Goodrich B, Gabry J, Ali I & Brilleman S. (2020). rstanarm: Bayesian applied regression modeling via Stan. R package version 2.21.1
  https://mc-stan.org/rstanarm.

Brilleman SL, Crowther MJ, Moreno-Betancur M, Buros Novik J & Wolfe R. Joint longitudinal and time-to-event models via Stan. StanCon 2018.
  10-12 Jan 2018. Pacific Grove, CA, USA. https://github.com/stan-dev/stancon_talks/
  
  